Structured Synchronous Reactive Programming for Game Development \\ \Large{Case Study: On Rewriting Pingus from C++ to \CEU}

### 2017-05-20 paper.tex
...

In Pingus, the game logic also accounts for almost half the size of the
codebase: 18.173 from 39.362 LoCs (46\%) spread in 272 files.
%
However, about half of the game logic relates to non-reactive code, such as
configurations and options, saved games and serialization, maps and levels
descriptions, string formatting, collision detection, graph algorithms, etc.
This part remains unchanged and relies on the seamless integration between \CEU
and C/C++~\cite{ceu.sensys13}.
%
Therefore, we rewrote 9.186 LoCs spread in 126 files%
\footnote{Game logic codebase: \url{https://github.com/sbgames-anon/paper-2017/tree/master/cpp/}}.
%
In order to only consider effective code in the analysis, we then removed all
headers, declarations, trivial getters \& setters, and other innocuous
statements, resulting in 70 implementation files with 4.135 dense LoCs
originally written in C++%
\footnote{``Effective'' C++ codebase: \url{https://github.com/sbgames-anon/paper-2017/tree/master/all/}}.
We did the same with the implementation in \CEU, resulting in 3.697 dense LoCs%
\footnote{``Effective'' \CEU codebase: \url{https://github.com/sbgames-anon/paper-2017/tree/master/all/}}.
%
Figure~\ref{tab.tree} summarizes the effective codebase in the two
implementations.

@@

In Pingus, a double click in the \emph{Armageddon button} at the bottom right
of the screen literally explodes all pingus.%
\footnote{Double click animation: \url{https://github.com/sbgames-anon/paper-2017/blob/master/README.md#1} }
Figure~\ref{lst.armageddon} compares the implementations in C++ and \CEU.

@@

The \emph{Bomber action} explodes the clicked pingu, throwing particles around
and also destroying the terrain under its radius.%
\footnote{Bomber action animation: \url{https://github.com/sbgames-anon/paper-2017/blob/master/README.md#2} }
%
We can model the explosion animation with a sequential state machine
(Figure~\ref{fig.bomber}) with actions associated to specific frames as
follows%
\footnote{State machine animation: \url{https://github.com/sbgames-anon/paper-2017/blob/master/README.md#3} }:
%
\begin{enumerate}
\item 0th frame:  plays a "Oh no!" sound.
\item 10th frame: plays a "Bomb!" sound.
\item 13th frame: throws particles, destroys the terrain, and shows an
                  explosion sprite.
\item Game tick:  hides the explosion sprite.
\item Last frame: kills the pingu.
\end{enumerate}

@@

The clickable \emph{blue dots} in the campaign world map transit to ambience
story screens%
\footnote{Story screen animation: \url{https://github.com/sbgames-anon/paper-2017/blob/master/README.md#4} }.
A story is composed of multiple pages and, inside each page, the words of the
story appear incrementally over time.
A first click in the button \code{>>>} fast forwards the words to show the full 
page.
A second click advances to the next page, until the story terminates.
If the page completes before a click (due to the time elapsing), a first click 
advances to the next page.
%
Figure~\ref{lst.story} compares the implementations in C++ and \CEU.

@@

A pingu is a dynamic entity created periodically and destroyed under certain
conditions, such as falling from a high altitude%
\footnote{Death of pingus animation: \url{https://github.com/sbgames-anon/paper-2017/blob/master/README.md#5} }.
%
Figure~\ref{lst.pingus} compares the implementations in C++ and \CEU.

@@

The \emph{mouse grab option} restricts the mouse movement to the game window
boundaries%
\footnote{Mouse grab animation: \url{https://github.com/sbgames-anon/paper-2017/blob/master/README.md#6} }.
The option can be set anywhere in the game by pressing \emph{Ctrl-G}.
In addition, the \emph{Options menu} has a check box to toggle the
\emph{mouse grab option} with mouse clicks while still responding to
\emph{Ctrl-G} presses.
%
Figure~\ref{lst.grab} compares the implementations in C++ and \CEU.

>>>>

In Pingus, the game logic also accounts for almost half the size of the
codebase: 18.173 from 39.362 LoCs (46\%) spread in 272 files.
%
However, about half of the game logic relates to non-reactive code, such as
configurations and options, saved games and serialization, maps and levels
descriptions, string formatting, collision detection, graph algorithms, etc.
This part remains unchanged and relies on the seamless integration between \CEU
and C/C++~\cite{ceu.sensys13}.
%
Therefore, we rewrote 9.186 LoCs spread in 126 files%
\footnote{Game logic codebase: \url{https://github.com/an000/p/tree/master/cpp/}}.
%
In order to only consider effective code in the analysis, we then removed all
headers, declarations, trivial getters \& setters, and other innocuous
statements, resulting in 70 implementation files with 4.135 dense LoCs
originally written in C++%
\footnote{``Effective'' C++ codebase: \url{https://github.com/an000/p/tree/master/all/}}.
We did the same with the implementation in \CEU, resulting in 3.697 dense LoCs%
\footnote{``Effective'' \CEU codebase: \url{https://github.com/an000/p/tree/master/all/}}.
%
Figure~\ref{tab.tree} summarizes the effective codebase in the two
implementations.

@@

In Pingus, a double click in the \emph{Armageddon button} at the bottom right
of the screen literally explodes all pingus.%
\footnote{Double click animation: \url{https://github.com/an000/p/blob/master/README.md#1} }
Figure~\ref{lst.armageddon} compares the implementations in C++ and \CEU.

@@

The \emph{Bomber action} explodes the clicked pingu, throwing particles around
and also destroying the terrain under its radius.%
\footnote{Bomber action animation: \url{https://github.com/an000/p/blob/master/README.md#2} }
%
We can model the explosion animation with a sequential state machine
(Figure~\ref{fig.bomber}) with actions associated to specific frames as
follows%
\footnote{State machine animation: \url{https://github.com/an000/p/blob/master/README.md#3} }:
%
\begin{enumerate}
\item 0th frame:  plays a "Oh no!" sound.
\item 10th frame: plays a "Bomb!" sound.
\item 13th frame: throws particles, destroys the terrain, and shows an
                  explosion sprite.
\item Game tick:  hides the explosion sprite.
\item Last frame: kills the pingu.
\end{enumerate}

@@

The clickable \emph{blue dots} in the campaign world map transit to ambience
story screens%
\footnote{Story screen animation: \url{https://github.com/an000/p/blob/master/README.md#4} }.
A story is composed of multiple pages and, inside each page, the words of the
story appear incrementally over time.
A first click in the button \code{>>>} fast forwards the words to show the full 
page.
A second click advances to the next page, until the story terminates.
If the page completes before a click (due to the time elapsing), a first click 
advances to the next page.
%
Figure~\ref{lst.story} compares the implementations in C++ and \CEU.

@@

A pingu is a dynamic entity created periodically and destroyed under certain
conditions, such as falling from a high altitude%
\footnote{Death of pingus animation: \url{https://github.com/an000/p/blob/master/README.md#5} }.
%
Figure~\ref{lst.pingus} compares the implementations in C++ and \CEU.

@@

The \emph{mouse grab option} restricts the mouse movement to the game window
boundaries%
\footnote{Mouse grab animation: \url{https://github.com/an000/p/blob/master/README.md#6} }.
The option can be set anywhere in the game by pressing \emph{Ctrl-G}.
In addition, the \emph{Options menu} has a check box to toggle the
\emph{mouse grab option} with mouse clicks while still responding to
\emph{Ctrl-G} presses.
%
Figure~\ref{lst.grab} compares the implementations in C++ and \CEU.
